A Theory for Data Migration of Information Systems

### 2022-09-17 2022 Migration/articleMigration.tex
Clarified disjoint union of datasets

	We can now define the notion of dataset.
	Let	$\concepts$ be a set of concepts,
	let $\isa$ be a partial order on $\concepts$,
	let $\rels$ be a set of relations,
	let $\atoms$ be a set of atoms,
	and let $\inst$ be a relation $\atoms\times\concepts$ that represents the instance relation between concepts and atoms,
\begin{definition}[dataset]
	\label{def:dataset}
	\item A dataset is a tuple $\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$ with
	\begin{eqnarray}
		\forall a\in\atoms\ \exists A\in\concepts&:&a\ \inst\ A\\
		\forall A,B\in\concepts, a\in\atoms&:&A\isa B\wedge a\ \inst\ A\ \Rightarrow\ a\ \inst\ B\\
		\forall r\in\rels&:&\src{r}\in\concepts\ \wedge\ \tgt{r}\in\concepts\\
		\forall\triple{a}{r}{b}\in\triples&:&r\in\rels\ \wedge\ a\ \inst\ \src{r}\ \wedge\ b\ \inst\ \tgt{r}
	\end{eqnarray}
\end{definition}

@@

\subsection{Data migration}
	To migrate information system $\infsys$ to $\infsys'$, we take the disjoint union $\infsys\sqcup\infsys'$ of both information systems
	and add rules to describe the migration.
	$\infsys$ is the existing system, which contains data to be preserved.
	$\infsys'$ is the new system, which contains no or hardly any data.
	It does, however, contain the concepts, relations, rules and roles of the new system.
\begin{figure}[bht]
	\begin{center}
	  \includegraphics[scale=.45]{Migration.drawio.png}
	\end{center}
\caption{Data migration}
\label{fig:event flow}
\end{figure}

	We use two datasets: $\dataset=\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$ and $\dataset'=\la\atoms',\concepts',\inst',\isa',\rels',\triples'\ra$.
	Before doing so, let us first define the disjoint union of two information systems.
\begin{definition}[disjoint union of datasets]
	\begin{eqnarray}
		\dataset\sqcup\dataset'&=&\la\atoms\sqcup\atoms',\ \concepts\sqcup\concepts',\ \inst\sqcup\inst',\ \isa\sqcup\isa',\ \rels\sqcup\rels',\ \triples\sqcup\triples'\ra\\
		\atoms\sqcup\atoms'&=&\{(a,0)|\ a\in\atoms\}\ \cup\ \{(a',1)|\ a'\in\atoms'\}\\
		\concepts\sqcup\concepts'&=&\{(c,0)|\ c\in\concepts\}\ \cup\ \{(c',1)|\ c'\in\concepts'\}\\
		\inst\sqcup\inst'&=&\begin{array}[t]{@{}l}
							  \{((a,0),(A,0))|\ (a,A)\in\inst\}\ \cup\\
			                  \{((a',1),(A',1))|\ (a',A')\in\inst'\}
							\end{array}\\
		\isa\sqcup\isa'&=&\begin{array}[t]{@{}l}
							  \{((A,0),(B,0))|\ A\isa B\}\ \cup\\
			                  \{((A',1),(B',1))|\ A'\isa' B'\}
							\end{array}\\
		\rels\sqcup\rels'&=&\{(r,0)|\ r\in\rels\}\ \cup\ \{(r',1)|\ r'\in\rels'\}\\
		\triples\sqcup\triples'&=&\begin{array}[t]{@{}l}
			                        \{\triple{(a,0)}{(r,0)}{(b,0)}|\ \triple{a}{r}{b}\in\triples\}\ \cup\\
			                        \{\triple{(a',1)}{(r',1)}{(b',1)}|\ \triple{a'}{r'}{b'}\in\triples'\}
								   \end{array}
	\end{eqnarray}
\end{definition}
\begin{definition}[$\doteq$]
	\begin{eqnarray}
		\declare{nm}{A}{B}\doteq\declare{nm'}{A'}{B'}&\Leftrightarrow&\id{nm}=\id{nm'}\wedge A=A'\wedge B=B'
	\end{eqnarray}
\end{definition}
\begin{definition}[migration dataset]
	\begin{eqnarray}
		&&\begin{array}[t]{@{}l}
			\{ {\tt ENFORCE\ }\declare{nm'}{A'}{B'}{\tt := }\ident{A'};\declare{nm}{A}{B};\ident{B'}\\
				|\ \begin{array}[t]{@{}l}
					\declare{nm}{A}{B}\in\rels\wedge\declare{nm'}{A'}{B'}\in\rels'\wedge\id{nm}=\id{nm'}\ \wedge\\
					\{\pair{A}{A'},\pair{B}{B'}\}\subseteq\isa\cup\isa'\cup\flip{\isa}\cup\flip{\isa'}\}
				   \end{array}
			  \end{array}\\
			  &\cup&\begin{array}[t]{@{}l}
				\{ {\tt CLASSIFY\ }(C',1){\tt\ IS\ }(C,0)|\ C\in\concepts,C'\in\concepts',C=C'\}\\
			  \end{array}
	\end{eqnarray}
\end{definition}
\begin{definition}[disjoint union of information systems]
	\begin{eqnarray}
		\infsys\sqcup\infsys'&=&\la\roles\sqcup\roles',\rules\sqcup\rules',\maintain\sqcup\maintain',\dataset\sqcup\dataset'\ra\\
		\roles\sqcup\roles'&=&\{(r,0)|\ r\in\roles\}\ \cup\ \{(r',1)|\ r'\in\roles'\}\\
		\rules\sqcup\rules'&=&\{(r,0)|\ r\in\rules\}\ \cup\ \{(r',1)|\ r'\in\rules'\}\\
		\maintain\sqcup\maintain'&=&\begin{array}[t]{@{}l}
							  \{((o,0),(u,0))|\ (o,u)\in\maintain\}\ \cup\\
			                  \{((o',1),(u',1))|\ (o',u')\in\maintain'\}
							\end{array}
	\end{eqnarray}
\end{definition}

>>>>

	We can now define the notion of dataset.
	Let	$\concepts$ be a set of concepts,
	let $\isa$ be a partial order on $\concepts$,
	let $\rels$ be a set of relations,
	let $\atoms$ be a set of atoms,
	and let $\inst$ be a relation $\atoms\times\concepts$ that represents the instance relation between concepts and atoms,
\begin{definition}[dataset]
	\label{def:dataset}
	\item A dataset is a tuple $\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$ with
	\begin{eqnarray}
		\forall a\in\atoms\ \exists A\in\concepts&:&a\ \inst\ A\\
		%\forall A,B\in\concepts, a\in\atoms&:&A\isa B\wedge a\ \inst\ A\ \Rightarrow\ a\ \inst\ B\\
		\forall \declare{n}{A}{B}\in\rels&:&A\in\concepts\ \wedge\ B\in\concepts\\
		\forall\triple{a}{\declare{n}{A}{B}}{b}\in\triples&:&\begin{array}[t]{@{}l}\declare{n}{A}{B}\in\rels\ \wedge\\ a\ (\inst \compose \kleenestar{\flip{\isa}})\ A\ \wedge\\ b\ (\inst \compose \kleenestar{\flip{\isa}})\ B\end{array}
	\end{eqnarray}
\end{definition}

@@

\subsection{Data migration}
	To migrate information system $\infsys$ to $\infsys'$, we take the disjoint union $\infsys\sqcup\infsys'$ of both information systems
	and add rules to describe the migration.
	$\infsys$ is the existing system, which contains data to be preserved.
	$\infsys'$ is the new system, which contains no or hardly any data.
	It does, however, contain the concepts, relations, rules and roles of the new system.
\begin{figure}[bht]
	\begin{center}
	  \includegraphics[scale=.45]{Migration.png}
	\end{center}
\caption{Data migration}
\label{fig:event flow}
\end{figure}

	We use two datasets: $\dataset=\la\atoms,\concepts,\inst,\isa,\rels,\triples\ra$ and $\dataset'=\la\atoms',\concepts',\inst',\isa',\rels',\triples'\ra$.
	Before doing so, let us first define the disjoint union of two information systems.
\begin{definition}[disjoint union of datasets]
	\begin{eqnarray}
		\dataset\sqcup\dataset'&=&\la\atoms\uplus\atoms',\ \concepts\uplus\concepts',\ \inst\uplus^2\inst',\ \isa\uplus^2\isa',\ \rels\uplus^3\rels',\ \triples\uplus^5\triples'\ra\\
		X\uplus Y&=&\{(x,0)|\ x\in X\}\ \cup\ \{(y,1)|\ y\in Y\}\\
		X\uplus^2 Y&=&\begin{array}[t]{@{}l}\{((x_1,0),(x_2,0))|\ (x_1,x_2)\in X\}\ \cup\\ \{((y_1,1),(y_2,1))|\ (y_1,y_2)\in Y\}\end{array}\\
		X\uplus^3 Y&=&\begin{array}[t]{@{}l}\{\declare{(n,0)}{(A,0)}{(B,0)}\mid\ \declare{n}{A}{B}\in X\}\ \cup\\ \{\declare{(n,1)}{(A,1)}{(B,1)}\mid\ \declare{n}{A}{B}\in Y\}\end{array}\\
		X\uplus^5 Y&=&\begin{array}[t]{@{}l}\{\triple{(a,0)}{\declare{(n,0)}{(A,0)}{(B,0)}}{(b,0)}\mid\ \declare{n}{A}{B}\in X\}\ \cup\\ \{\triple{(a,1)}{\declare{(n,1)}{(A,1)}{(B,1)}}{(b,1)}\mid\ \declare{n}{A}{B}\in Y\}\end{array}
	\end{eqnarray}
\end{definition}

\begin{definition}[migration dataset]
	\begin{eqnarray}
		&&\begin{array}[t]{@{}l}
			\{ {\tt ENFORCE\ }\declare{(nm',1)}{(A',1)}{(B',1)}\\\quad{\tt := }\ident{(A',1)};\declare{(nm,0)}{(A,0)}{(B,0)};\ident{(B',1)}\\
				\mid\ \begin{array}[t]{@{}l}
					\declare{nm}{A}{B}\in\rels\wedge\declare{nm'}{A'}{B'}\in\rels'\wedge\id{nm}=\id{nm'}\ \wedge\\
					\{\pair{A}{A'},\pair{B}{B'}\}\subseteq\isa\cup\isa'\cup\flip{\isa}\cup\flip{\isa'}\}
				   \end{array}
			  \end{array}\\
			  &\cup&\begin{array}[t]{@{}l}
				\{ {\tt CLASSIFY\ }(C',1){\tt\ IS\ }(C,0)|\ C\in\concepts,C'\in\concepts',C=C'\}\\
			  \end{array}
	\end{eqnarray}
\end{definition}
\begin{definition}[disjoint union of information systems]
	\begin{eqnarray}
		\infsys\sqcup\infsys'&=&\la\roles\uplus\roles',\rules\uplus^?\rules',\maintain\uplus^2\maintain',\dataset\sqcup\dataset'\ra
	\end{eqnarray}
\end{definition}
